Integrating Social Visual Analysis and Deployment with RCloud

### 2015-03-29 interviews.tex
trimming quotes a bit

To evaluate the effectiveness of RCloud, we interviewed 13 current and past
users of RCloud. Of these, 9 are data analysts, and 4 build tools for data
analysts and business needs.

@@

Raif says ``I've been sharing my notebooks in order for other people to see what
I've done. It's very convenient for that purpose, I would say.'' ``we haven't
been editing them together, but we've been just sharing, and I've been looking
at things that other people do. So there's no collaborative aspect of it, like
developing the code, but there's sharing aspects for other people to see and for
me to see other people's notebooks''

By default, all notebooks are publicly visible in RCloud, and notebooks can be
found through the notebook tree or search. However, the method for sharing that
users mentioned most is by sending links through email. Clint says, ``I can
share [my notebook] with people and copy it pretty readily. So if my supervisor
wants to see what I've done or QA it, I can just send her a link.''

Besides providing a way to present work to others, notebook sharing can also
provide a starting point. Clint says, ``I think the best part about it is how
easily you can share code [...] you can find a working example, rather than wearing
out Google and finding questionable examples that may or may not work.'' Raif
notes, if ``some person has done something similar, then you're able to just
edit that, and that's saved a lot of work time for me''

Taylor develops packages for analysts: ``I use it more for sharing code with
other people, and for doing tutorials for iotools or hmr'' or explaining issues
with data sources. ``A lot of times I'll share the code because I'm doing so
much package development. I want people to see how the package works, so I
clearly want them to see the code [...] I kind of write it just like I would write
GitHub Markdown, where you have little code snippets and text, but RCloud lets
me actually run the snippets'' and display the results.

Some users who tried out RCloud but are not able to use it for organizational
reasons, already miss the functionality. Shelly likes
``the concept of being able to create notebooks and share them.'' A ``wiki is not the
best way for communication results - it's kind of like writing a blog post with
very limited functionality [...] I have to save every
picture and post it as an image'' and ``I can't share all of the code because it
would just get crowded and wouldn't look right on a wiki.'' Bart says, ``if I could
make a folder on
RCloud and have Python notebooks and also Pig notebooks there, and execute them
from RCloud, that would be much better than my current thing, because that would
free me from manual documentation and version control and also telling people
where my code was. It would be just, hey, go look on RCloud, here's the stuff.''

@@

\subsection{Forking}
The ability to start work where someone else let off, by forking, has proved to
be a popular feature. In fact, almost twice as many (131) users have forked
someone else's notebook as have starred (75).  Clint: ``I use [forking] on my
own work quite a bit, and I use it on Research a ton. Honestly, it's one of the
handiest functions, because instead of having to find it, copy, paste it, you
just hit Fork, you rename it, and it's done. It's pretty amazing.''

Although we always intended forking to be used to improve others' code, we we
initially didn't support forking one's own notebooks, and this has proved very
useful. Rick says ``I fork my own notebooks because I'm going off and doing some
other analogous project, so I've sort of got interesting content that I've
already done in a previous analysis that I want to start from and then tweak to
match a new set of data.''

@@

Our users, especially those who build general purpose tools, see a need for better ways
to parameterize notebooks.  (Currently parameters can be added to the URL, but adding
user interface elements to do the same thing requires expertise.) Taylor complains that
a notebook might say `` `This is a report of the volume of all of
our feeds for this month', and someone would want to look at it for the next
month or the previous month, so they'd fork it to change the month.''

Jyothi calls it web-enabling the notebook: users ``can always fork the
notebooks and make changes, but I feel that if the owner of the notebook
web-enables it, then whoever is running it, it's easier to run. Instead of
forking it, if they can set options. It's probably more efficient, and also they
can [still] fork it if they want to.''

@@

Clint: ``instead of looking back and saying I've got a billion files here in
this subdirectory and I hope I've got them backed up, if they're on RCloud I
know they are [...] I don't have to worry about it if I drop my computer in the
bathtub, everythings gone, because I'm not great about uploading my own for
versioning''

Bart points out that the automatic versioning works well for dealing with the
minutiae of web development. ``I like the fact that it has a built-in editor, so
if you need to fix a typo in a link, or an extra line break or any of this other
nonsense, you can just switch to the edit view, pull up your asset, type
something, it's automatically saved, committed, everything. You don't have to go
back to your source code, change it, commit it to the repo, pull the repo to
your distribution version. None of that's necessary. A lot of web development is
fixing stupid little bugs, and you can do that instantly without a lot of
overhead, which is nice.''

@@

More selective ways to search will be needed as the corpus of notebooks
continues to grow. We are considering tagging notebooks filtering the notebook
tree by tag, measuring the quality of notebooks by their forks as well as stars,
and searching within a single user's notebooks.

\subsection{Web integration}
Integrating an analysis language into a web development environment is something
the tool developers really appreciate.  Horace: ``that also means the other guys
who want to do analytics on the data, they can first pull the data, do the
analytics on it, and then feed the viewer the data. So if anyone from the stats
group wants to insert some code in between the data pulling procedure, [they]
can always insert something in between. We were trying to make such an example
so that they don't have to rewrite the GUI part and they know how to get the
data source, once you get the data into RCloud, then you have a dataframe to
work with, and then they know how to produce another dataframe. We give them the
specification of the dataframes, we give them the data input, they do whatever
they want, they can reuse the data connection part and the GUI part. We can keep
changing these components and that's a very nice thing.''

Jyothi: ``Having an open session with R certainly helps too, because both of my
applications need to be doing something with the data, so having an open session
where I can run R commands or R functions without having to invoke an API or send
out a request and then wait for the request to come back to do the results,
that kind of seamless integration is extremely helpful in writing the
application.''

@@

\gordon{Not sure this belongs here}
Running anything on the Web can also be a source of frustration.  Poor network
links can cause web sockets to die, and Corey complains ``What frustrates me is
that if I do leave this window for what feels like 5 or 7 minutes, it
disconnects and I have to reconnect and run the whole notebook again.'' Raif
whines that ``being web-based, it's obviously slower than something you would be
running on your machine.'' For Byers: even the little lag of sending a command
to a server and receiving the result is intolerable

\subsection{Other workflows}
Although most of the analysts we spoke to use and appreciates the sharing
features, RCloud is less popular as a tool for exploratory data analysis and
development.

@@

Rick says, ``I find that the style of editing is clunky. I would never edit like that
if I'm doing ordinary work. I usually use vi.'' ``I mean it's nice to have it
saved, but there's this trade-off between it making it easier for me to present
something or to save something, and my ease of typing and correcting and things
in a plain old editor window.''

@@

@@

>>>>

To evaluate the effectiveness of RCloud, we interviewed 13 current and past
users of RCloud. Of these, 9 are data analysts and 4 build tools for data
analysts and business needs.

@@

Raif's usage is typical:
\begin{quote}
I've been sharing my notebooks in order for other people to see what
I've done. It's very convenient for that purpose. [...] We haven't
been editing them together, but we've been just sharing, and I've been looking
at things that other people do. So there's no collaborative aspect of it, like
developing the code, but there's sharing aspects for other people to see and for
me to see other people's notebooks.
\end{quote}

By default, all notebooks are publicly visible in RCloud, and notebooks can be
found through the notebook tree or search. However, the method for sharing that
users mentioned most is by sending links through email. Clint says, ``If my supervisor
wants to see what I've done or QA it, I can just send her a link.''

Besides providing a way to present work to others, notebook sharing can also
provide a starting point. Clint says, ``I think the best part about it is how
easily you can share code. [...] You can find a working example, rather than wearing
out Google and finding questionable examples that may or may not work.'' Raif
notes, if ``some person has done something similar, then you're able to just
edit that, and that's saved a lot of work time for me''

Taylor develops packages for analysts: ``I use it more for sharing code with
other people, and for doing tutorials for iotools or hmr'', the packages he
maintains. ``I want people to see how the package works, so I clearly want them
to see the code [...] I kind of write it just like I would write GitHub
Markdown, where you have little code snippets and text, but RCloud lets me
actually run the snippets'' and display the results. He also uses RCloud for
explaining issues with data sources.

Some users who tried out RCloud but are not able to use it for organizational
reasons, already miss the functionality. Shelly likes ``the concept of being
able to create notebooks and share them.'' A ``wiki is not the best way for
communicating results - it's kind of like writing a blog post with very limited
functionality [...] I have to save every picture and post it as an image'' and
``I can't share all of the code because it would just get crowded and wouldn't
look right on a wiki.'' Bart says, ``if I could make a folder on RCloud and have
Python notebooks and also Pig notebooks there, and execute them from RCloud,
that would be much better than my current thing, because that would free me from
manual documentation and version control and also telling people where my code
was. It would be just, hey, go look on RCloud, here's the stuff.''

@@

\subsection{Forking}
The ability to start work where someone else let off, by forking, has proved to
be a popular feature. In fact, almost twice as many (131) users have forked
someone else's notebook as have starred one (75).  Clint says, ``it's one of the
handiest functions, because instead of having to find it, copy, paste it, you
just hit Fork, you rename it, and it's done. It's pretty amazing.''

Although we always intended forking to be used to improve others' code, we we
initially didn't support forking one's own notebooks, and this has proved very
useful. Rick says ``I fork my own notebooks because I'm going off and doing some
other analogous project, so I've got interesting content that I've already done
in a previous analysis that I want to start from and then tweak to match a new
set of data.''

@@

Currently parameters can be added to the URL, but adding user interface elements
to do the same thing requires expertise. The tool-builders see a need for better
ways to parameterize notebooks. Taylor complains that a notebook might say ``
`This is a report of the volume of all of our feeds for this month', and someone
would want to look at it for the next month or the previous month, so they'd
fork it to change the month.''

Jyothi calls it ``web-enabling'' the notebook: users ``can always fork the
notebooks and make changes, but I feel that if the owner of the notebook
web-enables it [...] it's easier to run. Instead of
forking it, if they can set options, it's probably more efficient, and also they
can [still] fork it if they want to.''

@@

Automatic source control is also a popular feature. Clint says, ``instead of looking back and saying I've got a billion files here in
this subdirectory and I hope I've got them backed up, if they're on RCloud I
know they are. [...] I don't have to worry about it, if I drop my computer in the
bathtub, everythings gone, because I'm not great about uploading my own for
versioning.''

Bart points out that the automatic versioning works well for dealing with the
minutiae of web development:
\begin{quote}
I like the fact that it has a built-in editor, so
if you need to fix a typo in a link, or an extra line break or any of this other
nonsense, you can just switch to the edit view, pull up your asset, type
something, it's automatically saved, committed, everything. You don't have to go
back to your source code, change it, commit it to the repo, pull the repo to
your distribution version. None of that's necessary. A lot of web development is
fixing stupid little bugs, and you can do that instantly without a lot of
overhead, which is nice.
\end{quote}

@@

More selective ways to search will be needed as the volume of notebooks
continues to grow. We are considering tagging notebooks filtering the notebook
tree by tag, measuring the quality of notebooks by their forks as well as stars,
and searching within a single user's notebooks.

\subsection{Web integration}
Integrating an analysis language into a web development environment is something
the tool developers really appreciate.  The structure of Horace's visualization
notebook means
\begin{quote}
the other guys who want to do analytics on the data, they can first pull the
data, do the analytics on it, and then feed the viewer the data. So [...]
anyone from the stats group [...] can always insert something in between. We
were trying to make such an example so that they don't have to rewrite the GUI
part and they know how to get the data source. Once you get the data into
RCloud, then you have a dataframe to work with, and then they know how to
produce another dataframe.
\end{quote}

Integrating R also helps Jyothi's topic modeling tool, described in the last section.
\begin{quote}
Having an open session with R certainly helps too, because both of my
applications need to be doing something with the data, so having an open session
where I can run R commands or R functions without having to invoke an API or
send out a request and then wait for the request to come back to do the results,
that kind of seamless integration is extremely helpful in writing the
application.
\end{quote}

@@

\subsection{Other workflows}

Although most of the analysts we spoke to use and appreciates the sharing
features, RCloud is less popular as a tool for exploratory data analysis and
development.

@@

@@

But the web interface of RCloud doesn't cut it for him. ``I find that the style of editing is clunky. I would never edit like that
if I'm doing ordinary work. I usually use vi.'' ``I mean it's nice to have it
saved, but there's this trade-off between it making it easier for me to present
something or to save something, and my ease of typing and correcting and things
in a plain old editor window.''

@@

Running anything on the Web can also be a source of frustration.  Poor network
connections can cause web sockets to die, and Corey complains ``What frustrates me is
that if I do leave this window for what feels like 5 or 7 minutes, it
disconnects and I have to reconnect and run the whole notebook again.'' Raif
whines that ``being web-based, it's obviously slower than something you would be
running on your machine.'' For Byers: even the little lag of sending a command
to a server and receiving the result is intolerable
